---
title: "PL21/22 R Code"
output:
  html_document:
    df_print: paged
---

Firstly, we import some data for the sake of later usage.

```{r}
Tm <- c("Arsenal","Aston Villa","Brentford","Brighton and Hove Albion","Burnley"
        ,"Chelsea","Crystal Palace","Everton","Leeds United","Leicester City",
        "Liverpool","Manchester City","Manchester United","Newcastle United",
        "Norwich City","Southampton","Tottenham Hotspur","Watford","West Ham United", "Wolverhampton Wanderers")

TmAcr <- c("ARS","AVL","BRE","BHA","BUR","CHE","CRY","EVE","LEE","LEI","LIV",
           "MCI","MUN","NEW","NOR","SOU","TOT","WAT","WHU","WOL")

primcolours <- c("#ef0107","#7b003a","#e30613","#005daa","#630f35","#034694","#27409b","#274488","#ffffff","#0053a0","#dd0000", "#97c1e7","#b40509","#241f20","#fff200","#ed1a3b","#ffffff", "#fbee23","#7f0000","#fdb913")

secndcolours <- c("#ef0107","#7b003a","#ffffff","#ffffff", "#630f35","#034694","#c4122e","#274488","#ffffff","#0053a0","#dd0000","#97c1e7","#b40509","#ffffff","#fff200","#ffffff","#ffffff", "#fbee23","#7f0000","#fdb913")

bord <- c("#ffffff","#a3c5e9","#e30613","#005daa","#97d7f5","#ffffff","#a7a5a6","#ffffff","#1d428a","#fdbe11","#00b2a9","#1c2c5b","#fbe122","#241f20","#00a650","#ed1a3b","#132257","#11210c","#1bb1e7","#231f20")

```

Then we import the data sets:

```{r}
## replace the links to make this work
GoalMins <- read.csv('/Users/josiahwu29/Desktop/SummerResearchY1/GoalMins.csv')
GoalMinsFract <- read.csv('/Users/josiahwu29/Desktop/SummerResearchY1/GoalMinsFract.csv')
GoalMinsFractCumul <- read.csv('/Users/josiahwu29/Desktop/SummerResearchY1/GoalMinsFractCumul.csv')
FixturesByTeam <- read.csv('/Users/josiahwu29/Desktop/SummerResearchY1/FixturesByTeam.csv')
```

Now below is the code for computing the p-values:

```{r}
CountPerMatch <- function(t) {
return(GoalMins[GoalMins$Team==t,]$GoalsNum)
}
```

```{r}
ChiSqPoisMatch <- function(t) {
  v <- CountPerMatch(t)
  n <- length(v)
  ## Find q such that for groups with goals q => Expected Freq < 5.
  q <- qpois(1-(5/n),mean(v))
  goals <- c(0:q)
  ExpProb <- dpois(c(0:(q-1)),lambda=mean(v))
  ## Merge Groups with Expected Freq < 5
  ExpProb[q+1] <- (1 - ppois(q-1,mean(v)))
  Obse <- vector()
  for(k in 1:q) {
    Obse[k] <- length(which(v==k-1))
  }
  Obse[q+1] <- length(which(v>=q))

  return(chisq.test(Obse,p=ExpProb,rescale.p=TRUE))
}
```

```{r}
## Vector of p-values
ChiSqP <- c()
for(i in 1:20){
ChiSqP <- c(ChiSqP, ChiSqPoisMatch(Tm[i])$p.value)
}
```

```{r}
## Plot Bar Graph of p-values (Fiugre 2)

barplot(ChiSqP, names.arg = TmAcr, cex.names = 0.4, col = primcolours, ylim=c(0,1), xlab = 'Team', ylab = 'ChiSqTest p-value')

abline(h=0.9)
```

I made the 'guess' that the intensity function is a multiple of the PDF for Beta, based on my observations on the histograms...

```{r}
layout_matrix <- matrix(1:10,ncol=5,byrow=TRUE)
layout(layout_matrix)
## first 10 teams histogram
for(n in 1:10){
w <- GoalMinsFract[GoalMinsFract$Team==Tm[n],]$Decimal
hist(w, breaks=10, freq=FALSE, col = rep(c(primcolours[n],secndcolours[n]),5), border = bord[n], main=Tm[n], xlab = 'Proportion Of Match', ylab = 'Density')
}
```

```{r}
layout_matrix <- matrix(1:10,ncol=5,byrow=TRUE)
layout(layout_matrix)
## last 10 teams histogram
for(n in 11:20){
w <- GoalMinsFract[GoalMinsFract$Team==Tm[n],]$Decimal
hist(w, breaks=10, freq=FALSE, col = rep(c(primcolours[n],secndcolours[n]),5), border = bord[n], main=Tm[n], xlab = 'Proportion Of Match', ylab = 'Goals Scored')
}
```

and by using 'descdist' from fitdistrplus package.

```{r}
library(fitdistrplus)

layout_matrix2 <- c(1)
layout(layout_matrix2)
descdist(GoalMinsFract[GoalMinsFract$Team==Tm[1],]$Decimal)
```

Now we're ready to move on. We first input the intensity function and its integral:

```{r}
lambda <- function(theta, time){
    (theta[3]/beta(theta[1],theta[2])) * ((time%%1)**(theta[1]-1)) * ((1-(time%%1))**(theta[2]-1))
}

Lambda <- function(theta, time){
    theta[3]*pbeta((time%%1),theta[1],theta[2]) + theta[3]*floor(time)
}
```

as well as the log-likelihood. Note that here we're computing the NEGATIVE log-likelihood.

```{r}
nllh_ihpp <- function(theta, x, A){
    
    Lambda_term <- Lambda(theta, A[2]) - Lambda(theta, A[1])
    lambda_term <- - sum(log(lambda(theta, time = x))) 
    
    return(Lambda_term + lambda_term)
}
```

This is because the built-in function 'optim', by default, finds parameters that MINIMIZES the desired function. Hence we compute the negative log-likelihood as MIN(negative log-likelihood) \<=\> MAX(log-likelihood).

We can now write a function that finds the optimal parameters...

```{r}
opt <- function(t){
    ind <- which(Tm==t)
    opti <- optim(par = c(1, 1, 1),
                  fn = nllh_ihpp,
                  method = "L-BFGS-B",
                  lower = c(0, 0, 0),
                  upper = c(Inf, Inf, Inf),
                  hessian = TRUE,
                  A = c(0, 38),
                  x = GoalMinsFractCumul[GoalMinsFractCumul$Team==t,]$Decimal
    )
    return(opti)
}
```

and find such parameters for every team.

```{r}
theta1 <- vector()
theta2 <- vector()
theta3 <- vector()
Val <- vector()
Conv <- vector()
Team <- Tm

for(k in 1:20){
op <- opt(Tm[k])
theta1 <- append(theta1,op$par[1])
theta2 <- append(theta2,op$par[2])
theta3 <- append(theta3,op$par[3])
Val <- append(Val,op$value)
conv <- append(Conv,op$convergence)
}

ParaDF<- data.frame(Team, theta1, theta2, theta3, Val, conv)

ParaDF
```

We can finally write a function to simulate results for a single team.
```{r}
simulate_hpp_by_intervals <- function(t_min, t_max, lambda){
    event_times <- c()
    
    t_next <- t_min + rexp(n = 1, rate = lambda)
    
    while(t_next <= t_max){
        event_times <- c(event_times, t_next)
        t_next <- t_next + rexp(n = 1, rate = lambda)
    }
    
    return(event_times)
}
```

```{r}
simulate_opt <- function(Team){
    ti <- seq(0, 38, length.out = 5001)
    fitted_intensity <- lambda(theta = opt(Team)$par, time = ti)
    fitted_intensity <- fitted_intensity[is.finite(fitted_intensity)==TRUE]
    m <- 1.025*max(fitted_intensity)
    HPPResult <- simulate_hpp_by_intervals(0,38,m)
    
    Probs <- lambda(theta = opt(Team)$par, time = HPPResult)/m
    ZeroOne <- c()
    for(i in 1:length(HPPResult)){
        p = Probs[i]
        indicator = sample(c(0,1),size=1,prob=c((1-p),p))
        ZeroOne <- c(ZeroOne,c(indicator))
    }
    return(HPPResult[ZeroOne == 1])
}

```

```{r}
plot_sim_all <- function(s){
    set.seed(s)
    plot(NULL, xlim =c(0,38), ylim = c(0,4), yaxt='n', xlab='Match', ylab = 'Teams')
    for (i in 1:20){
        Team = Tm[i]
        W <- simulate_opt(Team)
        y_val = 4 - (i * 0.2)
        par(bg = "#c0c0c0")
        points(x = W, y = rep(y_val, length(W)), pch = 16, type = 'o', col=primcolours[i])
    }
}

plot_sim_all(2) ## Figure 3
```

And also generate a final League Table:

```{r}
ConvertGS <- function(v){
final <- c()
for(i in 1:38){
a <- length(v[v<=i & v>i-1])
final <- c(final, c(a))
}
return(final)}

ConvertSimGS <- function(t){
return(ConvertGS(simulate_opt(t)))
}

GenTable <- function(sd){
set.seed(sd)
v <- c()
for(i in 1:20){
v <- c(v,c(ConvertSimGS(Tm[i])))
}
points <- c(rep(0,20))
Wins <- c(rep(0,20))
Draws <- c(rep(0,20))
Losses <- c(rep(0,20))
GoalsFor <- c(rep(0,20))
GoalsAgainst <- c(rep(0,20))
df <- data.frame(
Team = FixturesByTeam$Team,
MatchID = FixturesByTeam$MatchID)
df['SimGoals'] = v
for(idx in 66342:66721){
dfidx <- df[df$MatchID == idx,]
g = dfidx$SimGoals
t = dfidx$Team
a_tm <- which(Tm==t[1])
b_tm <- which(Tm==t[2])
if(g[1]>g[2]){
points[a_tm] <- points[a_tm] + 3
Wins[a_tm] <- Wins[a_tm] + 1
Losses[b_tm] <- Losses[b_tm] + 1
}
if(g[2]>g[1]){
points[b_tm] <- points[b_tm] + 3
Wins[b_tm] <- Wins[b_tm] + 1
Losses[a_tm] <- Losses[a_tm] + 1
}
if(g[1]==g[2]){
points[a_tm] <- points[a_tm] + 1
points[b_tm] <- points[b_tm] + 1
Draws[a_tm] <- Draws[a_tm] + 1
Draws[b_tm] <- Draws[b_tm] + 1
}
GoalsFor[a_tm] <- GoalsFor[a_tm] + g[1]
GoalsFor[b_tm] <- GoalsFor[b_tm] + g[2]
GoalsAgainst[a_tm] <- GoalsAgainst[a_tm] + g[2]
GoalsAgainst[b_tm] <- GoalsAgainst[b_tm] + g[1]
}
GoalsDiff = GoalsFor - GoalsAgainst
df2 <- data.frame(Team = Tm, Points = points, W = Wins, D = Draws, L = Losses, GF = GoalsFor, GA = GoalsAgainst, GD = GoalsDiff)
df2 <- df2[order(df2$Points, decreasing=TRUE),]
return(df2)
}

try(GenTable(2))
```

At last, we can run the simulations 100 times, as follows:
```{r}
SamplekthPl <- function(k,size){
    kthPl <- c()
    i <- 1
    while(length(kthPl)<size){
        set.seed(i)
        try(kthPl <- c(kthPl, c(GenTable(i)$Team[k])))
        i <- i+1}
    return(table(kthPl))
  
}

SamplekthPl(1,100) ## error messages are normal; may differ from result in the poster.

```

